Revisions

### 2024-08-27 Manuscript/manuscript.qmd
re-render after a few minor revision text fixes

Project 2 will focus on the domain of function learning and in particular the issue of extrapolation. Function learning research examines how people acquire and generalize knowledge about continuous input-output relationships, and the factors influencing extrapolation to novel inputs following an initial learning phase. The domain of function learning has yielded influential computational models like the Associative Learning Model (ALM) and the Extrapolation-Association Model (EXAM) [ @busemeyerLearningFunctionalRelations1997], which have successfully accounted for human learning, interpolation, and extrapolation in numerous investigations[@deloshExtrapolationSineQua1997; @mcdanielConceptualBasisFunction2005; @mcdanielPredictingTransferPerformance2009]. However, the influence of training variability on function learning, particularly in visuomotor function learning tasks, remains relatively unexplored. Project 2 of this dissertation will address this gap by investigating how constant and varied training regimes affect learning, discrimination, and extrapolation in a novel visuomotor function learning task. We will leverage the ALM and EXAM models, fitted to individual participant data using advanced Bayesian techniques, to provide a detailed computational account of the observed empirical patterns.

@@

In Experiment 3, we sought to further explore the generality of the findings from the first two experiments by modifying the type of feedback provided during training. Specifically, we provided ordinal feedback instead of the continuous feedback used in the previous two experiments. Ordinal feedback provides learners with directional information about the results of their throw (e.g., above the target, below the target, or hitting the target) rather than precise numerical deviations. This form of feedback resembles many real-world learning scenarios, such as a coach instructing an athlete to perform a movement using "more force" or "less force", or a teacher providing letter grades rather than numeric scores. Although ordinal feedback provides less detailed information per trial, prior research has shown that less detailed feedback isn't necessarily detrimental to learning. For example, @cornwallEffectsCategoricalNumerical2022 manipulated whether participants received categorical (correct or incorrect) vs. numerical feedback (reward points ranging from 50-100). They found that the categorical condition produced superior learning, which they explained as arising from larger prediction errors. Although we do not make specific predictions about the ordinal condition in our study, it serves as a manipulation that might influence the learning process.

@@

It is important to note that while both projects utilize computational models, direct comparisons are complicated by the distinct purposes and structures of the models employed in each project. The IGAS model of Project 1 serves as a descriptive measurement model, capturing the similarity between training throws and testing conditions. In contrast, the ALM and EXAM models of Project 2 are full process models, capable of generating exact predictions for both learning and testing stages. The difference is also reflected in the interpretion of the generalization parameter ($c$) across the models of the two projects. In IGAS, $c$ moderates the similarity between executed throws and subsequent testing solutions, while in ALM and EXAM, $c$ governs the extent to which the perceived stimuli activate the input layer nodes. Despite these differences, insights from ALM/EXAM, particularly the role of zero-point knowledge, may offer potential explanations for the contrasting empirical results. Particularly, EXAM's reliance on zero-point knowledge in the simpler HTW task may explain why constant training was more effective in Project 2, while the lack of a clear zero-point reference in the more complex HTT task of Project 1 may have increased the value of varied training. This suggests that the benefits of variability depend critically on how task structure interacts with prior knowledge and the learner's capacity to leverage such knowledge for generalization.

>>>>

Project 2 will focus on the domain of function learning and in particular the issue of extrapolation. Function learning research examines how people acquire and generalize knowledge about continuous input-output relationships, and the factors influencing extrapolation to novel inputs following an initial learning phase. The domain of function learning has yielded influential computational models, including the Associative Learning Model (ALM) and the Extrapolation-Association Model (EXAM)[@busemeyerLearningFunctionalRelations1997], which have successfully accounted for human learning, interpolation, and extrapolation in numerous investigations[@deloshExtrapolationSineQua1997; @mcdanielConceptualBasisFunction2005; @mcdanielPredictingTransferPerformance2009]. However, the influence of training variability on function learning, particularly in visuomotor function learning tasks, remains relatively unexplored. Project 2 of this dissertation will address this gap by investigating how constant and varied training regimes affect learning, discrimination, and extrapolation in a novel visuomotor function learning task. We will leverage the ALM and EXAM models, fitted to individual participant data using advanced Bayesian techniques, to provide a detailed computational account of the observed empirical patterns.

@@

In Experiment 3, we sought to further explore the generality of the findings from the first two experiments by modifying the type of feedback provided during training. Specifically, we provided ordinal feedback instead of the continuous feedback used in the previous two experiments. Ordinal feedback provides learners with directional information about the results of their throw (e.g., above the target, below the target, or hitting the target) rather than precise numerical deviations. This form of feedback resembles many real-world learning scenarios, such as a coach instructing an athlete to perform a movement using "more force" or "less force", or a teacher providing letter grades rather than numeric scores. Although ordinal feedback provides less detailed information per trial, prior research has shown that less detailed feedback is not necessarily detrimental to learning. For example, @cornwallEffectsCategoricalNumerical2022 manipulated whether participants received categorical (correct or incorrect) vs. numerical feedback (reward points ranging from 50-100). They found that the categorical condition produced superior learning, which they explained as arising from larger prediction errors. While we do not make specific predictions about the ordinal condition, this manipulation allows us to explore how different types of feedback might interact with training variability to influence learning and generalization.

@@

It is important to note that while both projects utilize computational models, direct comparisons are complicated by the distinct purposes and structures of the models used in each project. The IGAS model of Project 1 serves as a descriptive measurement model, capturing the similarity between training throws and testing conditions. In contrast, the ALM and EXAM models of Project 2 are full process models, capable of generating exact predictions for both learning and testing stages. The difference is also reflected in the interpretion of the generalization parameter ($c$) across the models of the two projects. In IGAS, $c$ moderates the similarity between executed throws and subsequent testing solutions, while in ALM and EXAM, $c$ governs the extent to which the perceived stimuli activate the input layer nodes. Despite these differences, insights from ALM/EXAM, particularly the role of zero-point knowledge, may offer potential explanations for the contrasting empirical results. Particularly, EXAM's reliance on zero-point knowledge in the simpler HTW task may explain why constant training was more effective in Project 2, while the lack of a clear zero-point reference in the more complex HTT task of Project 1 may have increased the value of varied training. This suggests that the benefits of variability depend critically on how task structure interacts with prior knowledge and the learner's capacity to leverage such knowledge for generalization.

### 2024-08-27 revisions.qmd
re-render after a few minor revision text fixes

---
title: Revisions
toc: true
---

@@

<div>
It is important to note that while both projects utilize computational models, direct comparisons are complicated by the distinct purposes and structures of the models employed in each project. The IGAS model of Project 1 serves as a descriptive measurement model, capturing the similarity between training throws and testing conditions. In contrast, the ALM and EXAM models of Project 2 are full process models, capable of generating exact predictions for both learning and testing stages. The difference is also reflected in the interpretation of the generalization parameter ($c$) across the models of the two projects. In IGAS, $c$ moderates the similarity between executed throws and subsequent testing solutions, while in ALM and EXAM, $c$ governs the extent to which the perceived stimuli activate the input layer nodes. Despite these differences, insights from ALM/EXAM, particularly the role of zero-point knowledge, may offer potential explanations for the contrasting empirical results. Particularly, EXAM's reliance on zero-point knowledge in the simpler HTW task may explain why constant training was more effective in Project 2, while the lack of a clear zero-point reference in the more complex HTT task of Project 1 may have increased the value of varied training. This suggests that the benefits of variability depend critically on how task structure interacts with prior knowledge and the learner's capacity to leverage such knowledge for generalization.

@@

<details open class="addition-callout">
<summary>Addition [link](https://tegorman13.github.io/Dissertation/Manuscript/output/manuscript.html#experiment-3-38){target="_blank"}</summary>
<div>
> In Experiment 3, we sought to further explore the generality of the findings from the first two experiments by modifying the type of feedback provided during training. Specifically, we provided ordinal feedback instead of the continuous feedback used in the previous two experiments. Ordinal feedback provides learners with directional information about the results of their throw (e.g., above the target, below the target, or hitting the target) rather than precise numerical deviations. This form of feedback resembles many real-world learning scenarios, such as a coach instructing an athlete to perform a movement using "more force" or "less force", or a teacher providing letter grades rather than numeric scores. Although ordinal feedback provides less detailed information per trial, prior research has shown that less detailed feedback isn't necessarily detrimental to learning. For example, @cornwallEffectsCategoricalNumerical2022 manipulated whether participants received categorical (correct or incorrect) vs. numerical feedback (reward points ranging from 50-100). They found that the categorical condition produced superior learning, which they explained as arising from larger prediction errors. Although we do not make specific predictions about the ordinal condition in our study, it serves as a manipulation that might influence the learning process. 
</div>
</details>

@@

> Our results thus suggest that the benefits of variation cannot be explained by the varied-trained participants simply covering a broader range of the task space. Rather, the modeling suggests that varied participants also learn to adaptively tune their generalization function such that throwing locations generalize more broadly to one another than they do in the constant condition. A learning system could end up adopting a higher c value in the constant than variable training conditions by monitoring the trial-by-trial variability of the training items. The $c$ parameter would be adapted downwards when adjacent training items are dissimilar to each other and adapted upwards when adjacent training items are the same. In this fashion, contextually appropriate $c$ values could be empirically learned. This learning procedure would capture the insight that if a situation has a high amount variability, then the learner should be predisposed toward thinking that subsequent test items will also show considerable variability, in which case generalization gradients should be broad, as is achieved by low values for $c$. **@sakamotoTrackingVariabilityLearning2006 implemented a similar learning mechanism for updating the generalization paramater in an exemplar-based model (although in their model, a separate generalization parameter is assigned to each exemplar). In their experiment, participants were trained on a high variability and a low variability category, and the dynamically updated generalization parameter was necessary to account for broader generalione   zation observed around the high variability category when participants were tested with an ambiguous intermediary item. In a subsequent work [@sakamotoPuttingPsychologyBack2008], the same authors showed that a similar learning mechanism could account for the pattern wherein participants generalize more broadly around a category when the average distance between the category exemplars is larger (however the only model tested in this work was a prototype model).**
</details>

@@

<details open class="addition-callout">
<summary>Addition [link](https://tegorman13.github.io/Dissertation/Manuscript/output/manuscript.html#:~:text=Project%202%20will%20focus){target="_blank"}</summary>
<div>
Project 2 will focus on the domain of function learning and in particular the issue of extrapolation. Function learning research examines how people acquire and generalize knowledge about continuous input-output relationships, and the factors influencing extrapolation to novel inputs following an initial learning phase. The domain of function learning has yielded influential computational models like the Associative Learning Model (ALM) and the Extrapolation-Association Model (EXAM) [ @busemeyerLearningFunctionalRelations1997], which have successfully accounted for human learning, interpolation, and extrapolation in numerous investigations[@deloshExtrapolationSineQua1997; @mcdanielConceptualBasisFunction2005; @mcdanielPredictingTransferPerformance2009]. However, the influence of training variability on function learning, particularly in visuomotor function learning tasks, remains relatively unexplored. Project 2 of this dissertation will address this gap by investigating how constant and varied training regimes affect learning, discrimination, and extrapolation in a novel visuomotor function learning task. We will leverage the ALM and EXAM models, fitted to individual participant data using advanced Bayesian techniques, to provide a detailed computational account of the observed empirical patterns.
</div>
</details>

>>>>

---
title: Revisions
author:
- name: Thomas E. Gorman
  url: https://tegorman13.github.io/
page-layout: full
toc: true
format:
  html: default
  gfm: default
  docx:
    prefer-html: true
    output-file: "tg_revisions.docx"
    toc: false
---

@@

<div>
It is important to note that while both projects utilize computational models, direct comparisons are complicated by the distinct purposes and structures of the models used in each project. The IGAS model of Project 1 serves as a descriptive measurement model, capturing the similarity between training throws and testing conditions. In contrast, the ALM and EXAM models of Project 2 are full process models, capable of generating exact predictions for both learning and testing stages. The difference is also reflected in the interpretation of the generalization parameter ($c$) across the models of the two projects. In IGAS, $c$ moderates the similarity between executed throws and subsequent testing solutions, while in ALM and EXAM, $c$ governs the extent to which the perceived stimuli activate the input layer nodes. Despite these differences, insights from ALM/EXAM, particularly the role of zero-point knowledge, may offer potential explanations for the contrasting empirical results. Particularly, EXAM's reliance on zero-point knowledge in the simpler HTW task may explain why constant training was more effective in Project 2, while the lack of a clear zero-point reference in the more complex HTT task of Project 1 may have increased the value of varied training. This suggests that the benefits of variability depend critically on how task structure interacts with prior knowledge and the learner's capacity to leverage such knowledge for generalization.

@@

<details open class="addition-callout">
<summary>Addition [link](https://tegorman13.github.io/Dissertation/Manuscript/output/manuscript.html#experiment-3-38){target="_blank"}</summary>
<div>
> In Experiment 3, we sought to further explore the generality of the findings from the first two experiments by modifying the type of feedback provided during training. Specifically, we provided ordinal feedback instead of the continuous feedback used in the previous two experiments. Ordinal feedback provides learners with directional information about the results of their throw (e.g., above the target, below the target, or hitting the target) rather than precise numerical deviations. This form of feedback resembles many real-world learning scenarios, such as a coach instructing an athlete to perform a movement using "more force" or "less force", or a teacher providing letter grades rather than numeric scores. Although ordinal feedback provides less detailed information per trial, prior research has shown that less detailed feedback is not necessarily detrimental to learning. For example, @cornwallEffectsCategoricalNumerical2022 manipulated whether participants received categorical (correct or incorrect) vs. numerical feedback (reward points ranging from 50-100). They found that the categorical condition produced superior learning, which they explained as arising from larger prediction errors. While we do not make specific predictions about the ordinal condition, this manipulation allows us to explore how different types of feedback might interact with training variability to influence learning and generalization.
</div>
</details>

@@

> Our results thus suggest that the benefits of variation cannot be explained by the varied-trained participants simply covering a broader range of the task space. Rather, the modeling suggests that varied participants also learn to adaptively tune their generalization function such that throwing locations generalize more broadly to one another than they do in the constant condition. A learning system could end up adopting a higher c value in the constant than variable training conditions by monitoring the trial-by-trial variability of the training items. The $c$ parameter would be adapted downwards when adjacent training items are dissimilar to each other and adapted upwards when adjacent training items are the same. In this fashion, contextually appropriate $c$ values could be empirically learned. This learning procedure would capture the insight that if a situation has a high amount variability, then the learner should be predisposed toward thinking that subsequent test items will also show considerable variability, in which case generalization gradients should be broad, as is achieved by low values for $c$. **@sakamotoTrackingVariabilityLearning2006 implemented a similar learning mechanism for updating the generalization paramater in an exemplar-based model (although in their model, a separate generalization parameter is assigned to each exemplar). In their experiment, participants were trained on a high variability and a low variability category, and the dynamically updated generalization parameter was necessary to account for broader generalization observed around the high variability category when participants were tested with an ambiguous intermediary item. In a subsequent work [@sakamotoPuttingPsychologyBack2008], the same authors showed that a similar learning mechanism could account for the pattern wherein participants generalize more broadly around a category when the average distance between the category exemplars is larger (however the only model tested in this work was a prototype model).**
</details>

@@

<details open class="addition-callout">
<summary>Addition [link](https://tegorman13.github.io/Dissertation/Manuscript/output/manuscript.html#:~:text=Project%202%20will%20focus){target="_blank"}</summary>
<div>
Project 2 will focus on the domain of function learning and in particular the issue of extrapolation. Function learning research examines how people acquire and generalize knowledge about continuous input-output relationships, and the factors influencing extrapolation to novel inputs following an initial learning phase. The domain of function learning has yielded influential computational models, including the Associative Learning Model (ALM) and the Extrapolation-Association Model (EXAM)[@busemeyerLearningFunctionalRelations1997], which have successfully accounted for human learning, interpolation, and extrapolation in numerous investigations[@deloshExtrapolationSineQua1997; @mcdanielConceptualBasisFunction2005; @mcdanielPredictingTransferPerformance2009]. However, the influence of training variability on function learning, particularly in visuomotor function learning tasks, remains relatively unexplored. Project 2 of this dissertation will address this gap by investigating how constant and varied training regimes affect learning, discrimination, and extrapolation in a novel visuomotor function learning task. We will leverage the ALM and EXAM models, fitted to individual participant data using advanced Bayesian techniques, to provide a detailed computational account of the observed empirical patterns.
</div>
</details>
